ERNIE: A Knowledge Platform to Support Research Assessment

### 2018-01-11 LaTeX/Paper2/paper2.tex

Data mining coupled to network analysis has been successfully used as a digital methodology to study research collaborations and knowledge flow associated with drug development. To enable quantitative studies based on this approach, we have developed Enhanced Research Network Information Environment (ERNIE), a scalable cloud-based knowledge platform that integrates free data drawn from public sources as well as licensed data from commercially available ones. Analytical workflows in ERNIE are partially automated to enable expert input at critical stages. To facilitate adoption, reuse and extensibility, ERNIE is built with open source tools. and a modular design enables facile addition, deletion, or substitution of data sources. To demonstrate the capabilities of ERNIE, we report the results of seven case studies that span opioid addiction, pharmacogenomics, target discovery, behavioral interventions, and drug development. In these studies, we mine and analyze data from policy documents, regulatory approvals, research grants, bibliographic and patent databases, and clinical trials, to document collaborations and identify influential research accomplishments. ERNIE is presented as an accessible template for repositories that can be used to support expert qualitative assessments, while offering burden reduction through automation and access to integrated data.

@@

The risk of research evaluation being driven by data and metrics with insufficient consideration of balanced judgment has driven the formulation of  sound principles to advise practice~\cite{LeidenManifesto2015}. Data are still critical for evaluation, however, and utilities for capture, integration, curation, archival, and analysis that offer coverage and burden reduction can only assist the practice of research evaluation. Bibliographic data, often central to research evaluation, with the complementary use of administrative records such as government statistics and research funding are of considerable value~\cite{FedStat2017}. In this regard, Williams and colleagues have conducted network analysis of mined data from publicly available administrative, regulatory, and scientific resources to define relationships between scientific discoveries and major advances in medicine such as new drugs~\cite{Williams2015}. Extending this citation-centric approach, we have documented the scientific collaborations that extend across networks underlying the development of five independently developed therapeutics for cancer~\cite{Keserci2017}. Analysis of such networks supports new insights into collaboration, knowledge diffusion, and recursive learning \textit{(vide supra)}.

Cloud computing has also evolved to the point where scalable computing services and analytical methods can be made accessible to a broad user population and exerts a democratizing effect on research activities~\cite{Barga2011}. To take advantage of these advances in infrastructure and methodology, we have developed Enhanced Research Network Informatics Environment (ERNIE), a knowledge platform for research evaluation that integrates publicly available data as well as data from commercial sources. Focused on the theme of data mining and network analysis and to demonstrate the utility of ERNIE, we have used its data to conduct seven case studies. In these studies, we mine and analyze data from policy documents, regulatory approvals, research grants, bibliographic and patent databases, and clinical trials, to document collaborations and identify influential research accomplishments. The framework used to mine data and integrate it into networks for analysis incorporates expert input is simple, flexible, can be further automated, adapted to a range of subjects beyond drug development, and used to support individual researchers, groups, and research organizations engaged in research assessment. The workflows for these case studies incorporate expert knowledge into defining a set of core documents from which citations can be extracted, linked to other records in ERNIE, and analyzed. The first two case studies serve to validate ERNIE through reproducing the results of prior studies~\cite{Williams2015} on ivacaftor and ipilimumab, breakthrough drugs used to treat cystic fibrosis and melanoma. The remaining case studies relate to the current national crisis on substance abuse and concern two drugs used to treat opioid addiction~\cite{Blumberg1973, Cowan1977}, a microarray system for pharmacogenomic profiling~\cite{deLeon2006}, a target discovery system originating from enzyme fragment complementation assays that has been used in opioid receptor studies~\cite{Khanna1989}, and a behavioral intervention for substance abuse~\cite{Botvin1980}. The datasets that result from these studies are made available for use by other researchers and the workflows used to generate them are also archived to permit reproduction of our results, as well as review and modification of the methods used to generate them.

@@

\emph{Infrastructure} ERNIE runs on two IaaS CentOS 7.4 Linux servers in the Microsoft Azure cloud. One server, a standard D8s v3 (8 vCPUs, 32 Gb RAM) with 8 Tb of attached SSD storage, acts as a central hub for data related processes and hosts data in a PostgreSQL 9.6 relational database. The second server, a standard DS4 (8 vCPUs, 28 Gb) with 1 Tb of attached SSD storage also runs CentOS 7.4 and has the Apache Solr 7.1 search server installed. The design of ERNIE includes a PaaS Apache Spark cluster with 9 nodes  that is provisioned, using HDInsight, on demand and  is used for compute intensive tasks with Apache Sqoop, PySpark (Spark Python API), and data in Parquet format. Access to ERNIE servers is secured through SSH tunneling and two-factor authentication, with a unique time-based code acting as one of the factors. Interactive users have access to Python, SQL, R, Java, and standard Linux utilities. JupyterHub has also been installed to enable remote access to the Jupyter notebook server for multiple users.

\begin{figure}[!h]
%\begin{adjustwidth}{-0.5in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\scalebox{0.99}
{
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=.95\linewidth]{ERNIE_System_Context_2.pdf}
  \label{fig:sub1}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
 \includegraphics[width=.95\linewidth]{ERNIE_System_Context_1.pdf}
  \label{fig:sub2}
\end{subfigure}
}
\caption{{\bf System Context View} ERNIE runs on two IaaS CentOS 7.4 Linux servers in the Microsoft Azure cloud. \emph{Left Panel.} Users interactively access the environment through multifactor authentication (MFA). \emph{Right Panel.} ERNIE data is populated through automated custom ETL processes }
\label{fig: test}
%\end{adjustwidth}
\end{figure}

@@

\emph{Amplification.} Once unique identifiers to scientific literature are assembled in the seedset, cited and/or citing references in the Web of Science Core Collection are identified using the wos\_references table (see schema). Only those records with an entry in the wos\_publications table (see schema) are considered when counting citing or cited references. These are then linked to authors using Web of Science data. Wherever possible, UTs are mapped to pmids using a lookup table and then to grants using data linking publications to grant support from NIH ExPORTER.

\emph{Network analysis.} For network  analysis, 4 input files are used that are derived from the amplification stage (i) Article and Author edge list, a mapping of publications to authors. (ii) Article and first generation citations, a list of seedset and a single generation of cited or citing references derived from the amplification cycle. (iii) Publication and publication edge list, the citation network after amplification stored as source-target citation pairs. (iv) Publication and published year: A mapping of publications to their year of publication. Once these four files are generated, the network analyzer code generates 7 output files:

\begin{enumerate}
  \item Author scores based on PubMed and WoS. These are mappings of publications to their authors using either PubMed or Web of Science data. 
  \item Article scores based on PubMed and WoS: These one  contains the information of seed set and its first generation without inter citation information.
  \item Edge List of Articles for Visualization based on PubMed and WoS. These is the citation network of the publications after amplification.
  \item Descriptive statistics: This file contains  basic descriptive statistics about articles and authors such as counts, max citations, and number of lost nodes during the various mapping. 
\end{enumerate}

Sample metrics for Authors are number of articles, total citation and propagated in-degree rank (PIR) ~\cite{Williams2015}, and for articles we calculate the in network citation, weighted citation (propagated citation with depth = 2). In Each case study,  we calculate the scores as below:

Given the articles-to-articles citation relation, let matrix $G$ be a $n$x$n$ matrix where $n$ represent the number of articles in the network be a Matrix representation of the Network. We define 
the components $g_{ij}$ of the matrix $G$ as follows,

@@

where $u$ is a weight vector. Here, for different weight vector, we will get different measurements.
For example, for  $u=<1,1,1...,1>$,
\begin{equation*}
s(\mathfrak{p}_j) = [u*G]_{j} = \sum_k^n  u_k \cdot g_{jk}=\sum_k^n  1\cdot g_{jk}
\end{equation*}
which is just citation of the article $\mathfrak{p}_j$, denoted by $cit(\mathfrak{p}_j)$. There are various weight vector we could use such as weight based on the publication year, weight based on over all citation flow (~ Page rank), or we could have have normalized weight vector by it is average citation score. In these case studies we used one more weight vector to produce William et.all result. This new weight enable us not to count the number of citing papers, but also consider the citation based quality of citing papers. Considering the weight vector to be $u=<cit(\mathfrak{p}_1),cit(\mathfrak{p}_2),...,cit(\mathfrak{p}_n)>$, then weighted citation of a papers according to this weight vector will be

>>>>

Data mining coupled to network analysis has been successfully used as a digital methodology to study research collaborations and knowledge flow associated with drug development. To enable quantitative studies based on this approach, we have developed Enhanced Research Network Information Environment (ERNIE), a scalable cloud-based knowledge platform that integrates free data drawn from public sources as well as licensed data from commercially available ones. To facilitate adoption, reuse and extensibility, ERNIE is built with open source tools. and a modular design enables facile addition, deletion, or substitution of data sources. Additionally, analytical workflows in ERNIE are partially automated to enable expert input at critical stages. To demonstrate the capabilities of ERNIE, we report the results of seven case studies that span opioid addiction, pharmacogenomics, target discovery, behavioral interventions, and drug development. In these studies, we mine and analyze data from policy documents, regulatory approvals, research grants, bibliographic and patent databases, and clinical trials, to document collaborations and identify influential research accomplishments. ERNIE is presented as an accessible template for repositories that can be used to support expert qualitative assessments, while offering burden reduction through automation and access to integrated data.

@@

The risk of research evaluation being driven by data and metrics with insufficient consideration of balanced judgment has driven the formulation of  sound principles to advise practice~\cite{LeidenManifesto2015}. Supporting data for expert qualitative evaluation are still critical, however, and utilities for capture, integration, curation, and analysis that offer coverage and burden reduction will assist the practice of research evaluation.

Bibliographic data are usually central to research evaluation, and their value is complemented by the use of administrative records such as government statistics and research funding~\cite{FedStat2017}. 
Relevantly, Williams and colleagues have conducted network analysis of mined data from publicly available administrative, regulatory, and scientific resources to define relationships between scientific discoveries and major advances in medicine, such as new drugs~\cite{Williams2015}. 
Extending this citation-centric approach, we have previously documented the scientific collaborations that extend across networks underlying the development of five independently developed therapeutics for cancer~\cite{Keserci2017}. 
Analysis of such networks supports new insights into collaboration, knowledge diffusion, and recursive learning \textit{(vide supra)}.

As the scale of analysis increases, more robust computing resources become useful and sometimes necessary. Cloud computing has evolved to the point where scalable computing services and analytical methods are now available to a broad user population, and this availability exerts a democratizing effect on research activities~\cite{Barga2011}.

To take advantage of these advances in methodology and infrastructure, we have developed Enhanced Research Network Informatics Environment (ERNIE), a knowledge platform for research evaluation. ERNIE integrates publicly available data as well as data from commercial sources and the framework that is used to mine data and integrate it into networks for analysis is simple and flexible. The datasets that result from analysis can be made available for use by other researchers, and the workflows used to generate them are also archived to permit  review, and reproduction.

We demonstrate the utility of ERNIE in seven case studies. In these studies, we mine and analyze data from policy documents, regulatory approvals, research grants, bibliographic and patent databases, and clinical trials, to document collaborations and identify influential research accomplishments. The workflows for these case studies incorporate subject matter knowledge into defining a set of core documents from which citations can be extracted, linked to other records in ERNIE, and analyzed. The first two case studies serve to validate ERNIE through reproducing the results of prior studies on ivacaftor and ipilimumab~\cite{Williams2015}, breakthrough drugs used to treat cystic fibrosis and melanoma. The remaining case studies relate to the current national crisis on substance abuse and concern two drugs used to treat opioid addiction~\cite{Blumberg1973, Cowan1977}, a microarray system for pharmacogenomic profiling~\cite{deLeon2006}, a target discovery system originating from enzyme fragment complementation assays that has been used in opioid receptor studies~\cite{Khanna1989}, and a behavioral intervention for substance abuse~\cite{Botvin1980}.

ERNIE is presented as a simple extensible model for repositories that can be used to support expert qualitative assessments. Such repositories enable subject matter input into research evaluation while offering access to integrated data and burden reduction through automation. .

@@

\emph{Infrastructure} ERNIE runs on two IaaS CentOS 7.4 Linux servers in the Microsoft Azure cloud. One server, a standard D8s v3 (8 vCPUs, 32 Gb RAM) with 8 Tb of attached SSD storage, acts as a central hub for data related processes and hosts data in a PostgreSQL 9.6 relational database. The second server, a standard DS4 (8 vCPUs, 28 Gb) with 1 Tb of attached SSD storage also runs CentOS 7.4 and has the Apache Solr 7.1 search server installed. The design of ERNIE includes a PaaS Apache Spark cluster with 9 nodes  that is provisioned, using HDInsight, on demand and  is used for compute intensive tasks with Apache Sqoop, PySpark (Spark Python API), and data in Parquet format. Access to ERNIE servers is secured through SSH tunneling and two-factor authentication, with a unique time-based code acting as one of the factors. Interactive users have access to Python and Jupyter notebook, SQL, R, Java, and standard Linux utilities.

\begin{figure}[!h]
%\begin{adjustwidth}{-0.5in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
%\centering
\flushleft
\scalebox{0.99}
{
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=.95\linewidth]{ERNIE_System_Context_1.pdf}
  \label{fig:sub1}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
 % \centering
 \includegraphics[width=.95\linewidth]{ERNIE_System_Context_2.pdf}
  \label{fig:sub2}
\end{subfigure}
}
\caption{{\bf System Context View} ERNIE runs on two IaaS CentOS 7.4 Linux servers in the Microsoft Azure cloud. An Apache Spark cluster is provisioned on a PaaS basis for compute intensive tasks. \emph{Left Panel.} Users access the environment through multifactor authentication (MFA) and access the data server \emph{Right Panel.} ERNIE data is populated through automated custom ETL processes indexed by the Solr server for fast searches.}
\label{fig: test}
%\end{adjustwidth}
\end{figure}

@@

\emph{Amplification.} Once unique identifiers to scientific literature are assembled in the seedset, cited and/or citing references in the Web of Science Core Collection are identified using the wos\_references table (see schema). Only those records with an entry in the wos\_publications table (see schema) are considered when counting citing or cited references. These are then linked to authors using Web of Science data. Wherever possible, UTs are mapped to pmids using a lookup table and then to grants using data linking publications to grant support from NIH ExPORTER. As we have previously noted~\cite{Keserci2017}, the use of pmids for network construction results in significant loss of information since the intersection of bibliographic records between PubMed and Web of Science is relatively small.

\emph{Network analysis.} For network  analysis, 4 input files are used that are derived from the amplification stage (i) Article and Author edge list, a mapping of publications to authors. (ii) Article and first generation citations, a list of seedset and a single generation of cited or citing references derived from the amplification cycle. (iii) Publication and publication edge list, the citation network after amplification stored as source-target citation pairs. (iv) Publication and published year, a mapping of publications to their year of publication. Once these four files are generated, the network analyzer code generates 3 output files (i) Article scores based on weighted citation counts from within the network. (ii) A list of author scores based on article scores. (iii) Edge list of articles for Visualization. (iv) Descriptive statistics, this file contains descriptive statistics about articles and authors generated during network analysis. These output files are generated using both pmid-based or UT-based data, thus, [need to clarify].

For each article, we calculate in-network citations, as well as weighted citations (sum of article citations plus sum of citations of an article's first degree neighbors). For each author, we calculate the number of articles, total citations and propagated in-degree rank (PIR) ~\cite{Williams2015}.

Given the article-to-article relation, let matrix $G$ be a $n$x$n$ matrix where $n$ represents the number of articles in the network. We define 
the components $g_{ij}$ of the matrix $G$ as follows,

@@

where $u$ is a weight vector. Here, for different weight vector, we will get different measurements.
For example, for  $u=<1,1,1...,1>$,
\begin{equation*}
s(\mathfrak{p}_j) = [u*G]_{j} = \sum_k^n  u_k \cdot g_{jk}=\sum_k^n  1\cdot g_{jk}
\end{equation*}
which is citations of the article $\mathfrak{p}_j$, denoted by $cit(\mathfrak{p}_j)$. There are various weight vector we could use such as weight based on the publication year, weight based on over all citation flow (~ Page rank), or we could have have normalized weight vector by it is average citation score. In these case studies we used one more weight vector to produce William et.all result. This new weight enable us not to count the number of citing papers, but also consider the citation based quality of citing papers. Considering the weight vector to be $u=<cit(\mathfrak{p}_1),cit(\mathfrak{p}_2),...,cit(\mathfrak{p}_n)>$, then weighted citation of a papers according to this weight vector will be
